---
title: "V-transform paper: part 2"
author: "Alexander J. McNeil"
date: "`r Sys.Date()`"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(knitr.kable.NA = '')
```

This markdown script is the second of two scripts reproducing analyses in the paper "Modelling volatile time series using v-transforms and copulas" by Alexander J.~McNeil (2020).

## Libraries

We need the following:

```{r, echo = TRUE, warning = FALSE, message = FALSE}
library(tscopula) # main library
library(stats4) # for AIC
library(plyr) # rbinding dataframes
library(xtable) # nice tables
library(data.table) # for transpose data.frame
library(rugarch) # comparison of VaR estimate with GARCH
library(xts) # for plot.zoo
library(report) # for citations of reports
```
## Empirical v-transform

We will first look at the empirical v-transform for data on daily Bitcoin log-returns for the period 2016-2019.

```{r, fig.show='hold', dev.args =list(pointsize=9)}
load("BTCUSD1619.RData")
BTCUSD <- BTCUSD1619$`BTCUSD=X.Close` # closing values
X <- (diff(log(BTCUSD))[-1]) * 100
U <- strank(X)
V <- strank(abs(X))
plot(as.numeric(U), as.numeric(V), xlab= "u", ylab = "v", type="p")
plot(acf(qnorm(V), plot = FALSE), ylab="ACF of z", main= " ")
```

## Fitting VT-ARMA copula models to data

We start with a Markovian AR(1) model and the linear v-transform. Optimization is by Nelder-Mead.


```{r, echo = TRUE}
tsoptions <- list(hessian = TRUE, method = "Nelder-Mead", avoidzero= FALSE)

modspec <- vtscopula(armacopula(list(ar = 0.5)), Vlinear(0.45))
vt1arma10 <- fit(modspec, U, tsoptions = tsoptions)
```

### ARMA(1,1) models

The next set of models are ARMA(1,1) models with an increasingly more complicated v-transform. The starting parameters are always set to be close to the optimum from the previous fit.

```{r, echo = TRUE}
modspec <- vtscopula(armacopula(list(ar = 0.95, ma = -0.85)), Vlinear(0.45))
vt1arma11 <- fit(modspec, U, tsoptions = tsoptions)
modspec <- vtscopula(armacopula(list(ar = 0.96, ma = -0.84)), V2p(delta=0.42, kappa=1))
vt2arma11 <- fit(modspec, U, tsoptions = tsoptions)
modspec <- vtscopula(armacopula(list(ar = 0.9615, ma = -0.836)), V3p(delta=0.462, kappa=1, xi = 1))
vt3arma11 <- fit(modspec, U, tsoptions = tsoptions)
```


### Process results

Results are summarized in a table.

```{r, echo = TRUE}
models <- c("vt1arma10", "vt1arma11", "vt2arma11", 
            "vt3arma11")
process_results <- function(){
  allresults <- list()
  for (i in 1:length(models)){
    results <- eval(parse(text = models[i]))
    copmod <- results@tscopula
    V <- vtrans(copmod@Vtransform, U, correction = TRUE)
    resid <- kfilter(copmod@Vcopula, V)[, "resid"]
    SW <- shapiro.test(as.numeric(resid))$p.value
    pars <- coef(results)
    ses <- safe_ses(results@fit$hessian)
    names(ses) <- names(pars)
    pars <- as.list(pars)
    ses <- as.list(ses)
    df1 <- data.frame(name = models[i], pars , SW,
                      logLik = as.numeric(logLik(results)), AIC = AIC(results))
    df2 <- data.frame(ses)
    allresults[[2*(i-1)+1]] <- df1
    allresults[[2*(i-1)+2]] <- df2
  }
  allresults
}
allresults <- process_results()
tableout <- do.call(rbind.fill, allresults)
tableout <- tableout[,c("name", "ar1", "ma1", "delta", "kappa", "xi","SW","logLik","AIC")]
colnames(tableout) <- c("Model", "$\\alpha_1$", "$\\beta_1$", "$\\delta$", "$\\kappa$", "$\\xi$", "SW", "$L$", "AIC") 
knitr::kable(tableout, digits = 3)
```

According to AIC the most effective model is the VT-ARMA(1,1) model with 2-parameter v-transform. The following graphs show various diagnostics for the fit.

```{r, echo=TRUE, fig.show='hold', dev.args =list(pointsize=9)}
plot(vt2arma11, type=3, bw = TRUE)
plot(vt2arma11, type=4, bw = TRUE)
plot(vt2arma11, type=5, bw = TRUE)
plot(vt2arma11, type=2, bw = TRUE)
```

## Adding a parametric marginal distribution

We now add parametric margins to the best VT-ARMA(1,1) copula model. 

```{r, fullmodels}
marg_st <- fit(margin("st"), X)
marg_laplace <- fit(margin("laplace", 
                           pars = c(mu = 0.2, scale = 1.5)), X)
marg_dweibull <- fit(margin("doubleweibull", 
                            pars = c(mu = 0.2, shape =0.8, scale = 2.7)), X)

vtcop <- vt2arma11

mod_st <- tscm(vtcop, margin = marg_st)
mod_st <- fit(mod_st, X, method = "full", tsoptions = tsoptions)
mod_laplace <- tscm(vtcop, margin = marg_laplace)
mod_laplace <- fit(mod_laplace, X, method = "full", tsoptions = tsoptions)
mod_dweibull <- tscm(vtcop, margin = marg_dweibull)
mod_dweibull <- fit(mod_dweibull, as.numeric(X), method = "full", tsoptions = tsoptions)
```

These are the results.

```{R, echo = TRUE}
models <- c("mod_st", "mod_laplace", "mod_dweibull")
names(mod_st@margin@pars) <- c("shape", "mu", "sigma")
names(mod_laplace@margin@pars) <- c("mu", "sigma")
names(mod_dweibull@margin@pars) <- c("mu", "shape","sigma")
fullresults <- process_results()
tableout <- do.call(rbind.fill, fullresults)
tableout <- tableout[,c("ar1","ma1","delta","kappa","shape",
                        "mu","sigma","SW","logLik","AIC")]
tableout <- transpose(tableout)
rownames(tableout) <- c("$\\alpha_1$",  "$\\beta_1$", "$\\delta$", "$\\kappa$", "$\\eta$", "$\\mu$", "$\\sigma$", 
                        "SW", "$L$", "AIC")
colnames(tableout) <- c(models[1], "s.e", models[2], "s.e.", models[3], "s.e.")
knitr::kable(tableout, digits = 3)
```

Graphical analyses of fitted double Weibull.

```{R, echo=TRUE, fig.show='hold', dev.args =list(pointsize=9)}
names(mod_dweibull@margin@pars) <- c("mu", "shape","scale")
plot(mod_dweibull, type =2, bw = TRUE)
plot(mod_dweibull, type =3, bw = TRUE)
plot(mod_dweibull, type =6, bw = TRUE)
plot(mod_dweibull, type =4, bw = TRUE)
```

Some implications of the fitted model: we estimate the change point value $\mu_T$; we also calculate the volatility response for log-return values of -10 and +10.

```{R, echo=TRUE, fig.show='hold', dev.args =list(pointsize=9)}
qmarg(mod_dweibull@margin, mod_dweibull@fit$par["vt.delta"])
qnorm(vtrans(mod_dweibull@tscopula@Vtransform,pmarg(mod_dweibull@margin, c(-10,10))))
```

## Comparison with GARCH models

We now compare the VT-ARMA models with GARCH(1,1) models using Student and GED (generalized error distribution) innovations. Note that skewed innovations do not significantly improve the fit of either GARCH model.

```{R, echo = TRUE}
garchmod_ged <- ugarchspec(variance.model = list(
  model="sGARCH", 
  garchOrder=c(1,1)),
  mean.model=list(armaOrder=c(0,0),include.mean=TRUE),
  distribution.model = "ged")
mod_garch_ged <- ugarchfit(garchmod_ged, X)

garchmod_std <- ugarchspec(variance.model = list(
  model="sGARCH", 
  garchOrder=c(1,1)),
  mean.model=list(armaOrder=c(0,0),include.mean=TRUE),
  distribution.model = "std")
mod_garch_std <- ugarchfit(garchmod_std, X)

knitr::kable(AIC(mod_st, mod_laplace, mod_dweibull, 
    mod_garch_std, mod_garch_ged), digits=3)
```
## VaR calculation

The plot shows a comparison of the 95\% VaR estimated by the best VT-ARMA model (red) and the best GARCH model (black).

```{R, echo=TRUE, fig.show='hold', fig.width = 6, fig.height = 4, dev.args =list(pointsize=9)}
alpha <- 0.05
VaR <- quantile(mod_dweibull, alpha)
VaR_alt <- quantile(mod_garch_ged, alpha)
plot.zoo(cbind(X, VaR_alt, VaR), 
         plot.type="single", 
         col=c("grey","black","red"), 
         xlab="", ylab="", type=c("h", "l", "l"))
```

## Out-of-sample forecasting

The final section requires some patience. We first set up the data for forecasting.

```{R, echo = TRUE}
load("BTCUSD1219.RData")
BTCUSD1219 <- BTCUSD1219$`BTCUSD=X.Close` # closing values
X2 <- (diff(log(BTCUSD1219))[-1]) * 100
X2 <- X2[-(1:43)]
n0 <- 1000 # length of moving window
nforecast <- length(X2) - n0 # number of forecasts (period 2016-19)
```

We use the best VT-ARMA model for forecasting.

```{R, echo = TRUE}
tsoptions <- list(hessian = FALSE, method = "Nelder-Mead", avoidzero= FALSE)
VaR <- matrix(NA,ncol=2, nrow = nforecast)
for (i in 1:nforecast){
  fitdata <- as.numeric(X2[i:(i+n0-1)])
  fit <- fit(mod_dweibull, fitdata, tsoptions = tsoptions)
  VaR[i,1] <- quantile(fit, 0.05,last = TRUE)
  VaR[i,2] <- quantile(fit, 0.01, last = TRUE)
}
```

We also use the best GARCH model for forecasting.

```{R, echo = TRUE}
roll <- ugarchroll(garchmod_ged, X2, n.ahead = 1, 
                   forecast.length = nforecast, 
                   refit.every = 1, 
                   refit.window = "moving",
                   VaR.alpha = c(0.05,0.01), keep.coef = F)
```

We compare results and plot them as before.

```{R, echo = TRUE, fig.show='hold', fig.width = 6, fig.height = 4, dev.args =list(pointsize=9)}
sum(roll@forecast$VaR$realized < roll@forecast$VaR$`alpha(5%)`)
sum(roll@forecast$VaR$realized < VaR[,1])
nforecast*0.05

sum(roll@forecast$VaR$realized < roll@forecast$VaR$`alpha(1%)`)
sum(roll@forecast$VaR$realized < VaR[,2])
nforecast*0.01

Xforecast <- X2[-(1:n0)]
VaR_GARCH <- roll@forecast$VaR$`alpha(5%)`
attributes(VaR_GARCH) <- attributes(Xforecast)
VaR_VTARMA <- VaR[,1]
attributes(VaR_VTARMA) <- attributes(Xforecast)
plot.zoo(cbind(Xforecast, VaR_GARCH, VaR_VTARMA), 
         plot.type="single", 
         col=c("grey","black","red"), 
         xlab="", ylab="", type=c("h", "l", "l"))
```


```{R}
cite_packages(sessionInfo())
```